XXX in a Tracing JIT Compiler for Efficient Dynamic Languages

### 2011-03-24 talk/icooolps2011/paper.tex
expand intro

long dream of PE to not have to implement a compiler
new life by meta-tracers, spur, pypy, other things
trace execution of object model and get semantics right, not have to compile
object model

in spur and pypy, this can be improved by the interpreter author by adding
hints, which influence tracer and optimizer

pypy's hints go further than spurs, in this paper we present two very important
ones and show how classical implementation techniques of dynlangs can be
expressed by them

@@

One disadvantage of tracing JITs which makes them not directly applicable to Pypy,

PyPy's JIT is a meta-tracer \cite{bolz_tracing_2009}. Since we want to re-use
our tracer for a variety of languages, we
don't trace the execution of the user program, but instead trace the execution
of the \emph{interpreter} that is running the program. This means that the traces
don't contain the bytecodes of the language in question, but RPython-level
operations that the interpreter did to execute the program.

>>>>

It has long been an objective of the partial evaluation community to
automatically produce compilers from interpreters. There has been a recent
renaissance of this idea using the different technique of tracing just-in-time
compilers. A number of projects have attempted this approach. SPUR \cite{XXX}
is a tracing JIT for .NET together with a JavaScript implementation in C\#. XXX

All these projects have in common that they trace an implementation language
which is then used to implement an object model of a dynamic language. The
tracer then traces through this object model, which makes the object model
transparent to the tracer and its optimizations. Therefore the semantics of the
dynamic language does not have to be replicated in the JIT. We call this
approach \emph{meta-tracing}. We will give an introduction to the PyPy project
and to meta-tracing in Section~\ref{sec:Background}.

Another commonality of all these approaches is that they require some
annotations (or hints) in the dynamic language implementation to guide the
meta-tracer. SPUR and PyPy both provide the interpreter author with more
elaborate hints to influence the meta-tracer and its optimizer.

PyPy's hints go even further than SPUR's in that they provide the interpreter
author with a flexible toolset to make her implementation extremely efficient.
In this paper we present the two most prominent ones and show how classical
implementation techniques of dynamic languages can be expressed with them.

The contributions of this paper are:
\begin{itemize}
 \item A hint to introduce arbitrary constants into the trace.
 \item A way to define new pure operations which the optimizer then recognizes.
 \item A worked-out example of a simple object model of a dynamic language and how it can be improved using these hints.
\end{itemize}

@@

One disadvantage of tracing JITs which makes them not directly applicable to
PyPy is that they encode the language semantics. Since PyPy wants to be a
general framework, we want to reuse our tracer for different languages.
Therefore PyPy's JIT is a meta-tracer \cite{bolz_tracing_2009}. It does not
trace the execution of the user program, but instead traces the execution of
the \emph{interpreter} that is running the program. This means that the traces
it produces don't contain the bytecodes of the language in question, but
RPython-level operations that the interpreter did to execute the program.
